---
title: "Project 1"
author: "Brandy Carr"
format: 
  html:
    self-contained: true
    echo: true
    message: false
    warning: false
    error: true # IMPORTANT NOTE - this makes the file render even if there is an error. it is imperative that you check your .html document *before submission* to make sure that it has all results in it.
editor: source
---

**This week you will be analyzing data from the Jackson Heart Study (JHS). You can find the data on Canvas. For full credit, you must include all code chunks and R output backing up your responses.**

**0. Import the JHS data; you can download it from Canvas. You need to research how to read a SAS data file into R (hint: look into the `haven` package).**

```{r}

```

Insert your answer here :) 

**1a. Model systolic blood pressure (*sbp*; mmHg) as a function of age (*age*; years), education (*HSgrad*; 0=no, 1=yes), and body mass index (*BMI*; kg/m<sup>2</sup>). Remember to report the resulting model.**

```{r}

```

Insert your answer here :) 

**1b. Construct the 95% confidence intervals for the regression coefficients. Remember to state the resulting confidence intervals.**

```{r}

```

Insert your answer here :) 

**1c. Is this a significant regression line? Test at the $\alpha=0.05$ level. Remember to typeset the results.**

```{r}

```

Insert your answer here :) 

**1d. Which predictors, if any, are significant predictors of systolic blood pressure? Test at the $\alpha=0.05$ level. Remember to typeset the results.**

```{r}

```

Insert your answer here :) 

**1e. Provide brief interpretations for the slopes of the predictors.**

Insert your answer here :)

**1f. Construct an appropriate data visualization to help with explaining the model results. Systolic blood pressure should be on the y-axis, age should be on the x-axis. You choose what lines to create.**

```{r}

```

**1g. How many suspected outliers exist? You must justify your answer statistically.**

```{r}

```

Insert your answer here :) 

**1h. How many suspected influential/leverage points exist? You must justify your answer statistically.**

```{r}

```

Insert your answer here :) 

**1i. Is multicollinearity a problem in this model? You must justify your answer statistically.**

```{r}

```

Insert your answer here :) 

**1k. Assess the assumptions on the linear model. Remember to draw your conclusion with appropriate justification.**

```{r}

```

Insert your answer here :) 

**2a. Model systolic blood pressure (*sbp*; mmHg) as a function of age (*age*; years), education (*HSgrad*; 0=no, 1=yes), and health status as defined by body mass index (*BMI3cat*; 0=poor health, 1=intermediate health, 2=ideal health). Remember to report the resulting model.**

```{r}

```

Insert your answer here :) 

**2b. Construct the 95% confidence intervals for the regression coefficients. Remember to state the resulting confidence intervals.**

```{r}

```

Insert your answer here :) 

**2c. Which predictors, if any, are significant predictors of systolic blood pressure? Test at the $\alpha=0.05$ level. Remember to typeset the results.**

```{r}

```

Insert your answer here :) 

**2d. Provide brief interpretations for the slopes of the predictors.**

Insert your answer here :)

**2e. Construct an appropriate data visualization to help with explaining the model results. Systolic blood pressure should be on the y-axis, age should be on the x-axis. You choose what lines to create.**

```{r}

```

**3. Required for graduate students / extra credit for undergraduate students: Write a paragraph to summarize the above analysis, written such that a non-quantitative person could understand. Include if you recommend using BMI as a continuous variable or a categorical variable. For full credit, you must explain your reasoning.**

Insert your answer here :) 
