---
title: "Project 1"
subtitle: "STA 6232 Stats for Data Science II"
author: "Brandy Carr"
format: 
  html:
    fontcolor: "#388061"
    backgroundcolor: "#F9F8F1"
    toc: true
    number-sections: true
    number-depth: 2
    shift-heading-level-by: -1
    error: true
    smooth-scroll: true
execute: 
  warning: false
  message: false
editor: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(formattable)

# COLOR PALETTE
pal <- colorspace::terrain_hcl(7, h=c(150,0), c=c(45,45), l=c(60,75), power=c(1.5,1.5))
```

**This week you will be analyzing data from the Jackson Heart Study (JHS). You can find the data on Canvas. For full credit, you must include all code chunks and R output backing up your responses.**

</br>

### 0 {.unnumbered}

Import the JHS data; you can download it from Canvas. You need to research how to read a SAS data file into R (hint: look into the `haven` package).

::: {.panel-tabset}

## Code

```{r}
#| eval: false
JHS <- haven::read_sas("analysis1.sas7bdat")
DT::datatable(head(JHS, n=10), filter="top") # first 10 rows of data
```

## Output

```{r}
#| echo: false
JHS <- haven::read_sas("analysis1.sas7bdat")
DT::datatable(head(JHS, n=10), filter="top") # first 10 rows of data
```

:::

</br>

###

#### 1a

Model systolic blood pressure (*sbp*; mmHg) as a function of age (*age*; years), education (*HSgrad*; 0=no, 1=yes), and body mass index (*BMI*; kg/m<sup>2</sup>). Remember to report the resulting model.

::: {.panel-tabset}

## Code

```{r}
#| eval: false
# Checked for incomplete/missing cases
table(complete.cases(JHS %>% select(sbp, age, HSgrad, BMI)))

# removed incomplete cases
JHS1 <- JHS %>% select(sbp, age, HSgrad, BMI) %>% na.omit()

m1 <- glm(sbp ~ ., data=JHS1, family="gaussian")
summary(m1)
```

## Output

```{r}
#| echo: false
# Checked for incomplete/missing cases
table(complete.cases(JHS %>% select(sbp, age, HSgrad, BMI)))

# removed incomplete cases
JHS1 <- JHS %>% select(sbp, age, HSgrad, BMI) %>% na.omit()

m1 <- glm(sbp ~ ., data=JHS1, family="gaussian")
summary(m1)
```

## Results

*11 rows containing missing data were removed*

$\hat{sbp} = 95.04 + 0.44(age) - 0.84(HSgrad) + 0.24(BMI)$

:::

</br>

#### 1b

Construct the 95% confidence intervals for the regression coefficients. Remember to state the resulting confidence intervals.

::: {.panel-tabset}

## Code

```{r}
#| eval: false
confint(m1)
```

## Output

```{r}
#| echo: false
confint(m1)
```

## Results

- 95% CI for $\beta_{age}$ is (0.39, 0.49)

- 95% CI for $\beta_{HSgrad}$ is (-2.51, 0.83)

- 95% CI for $\beta_{BMI}$ is (0.16, 0.32)

:::

</br>

#### 1c

Is this a significant regression line? Test at the $\alpha=0.05$ level. Remember to typeset the results.

::: {.panel-tabset}

## Code

```{r}
#| eval: false
m0 <- glm(sbp ~ 1, data=JHS1, family="gaussian") # intercept only model
anova(m0, m1, test="F")
```

## Output

```{r}
#| echo: false
m0 <- glm(sbp ~ 1, data=JHS1, family="gaussian") # intercept only model
anova(m0, m1, test="F")
```

## Results

**Hypotheses:**

- $H_{0}$ : $\beta_{age} = \beta_{HSgrad} = \beta_{BMI} = 0$

- $H_{1}$ : at least one $\beta_{i} \ne 0$

**Test Statistic:** 

- $F_{0} = 122.13$

**P-Value:** 

- $p < 0.001$

**Rejection Region:** 

- $\alpha = 0.05$

- Reject $H_{0}$ if $p < \alpha$

**Conclusion:** 

- Reject $H_{0}$ 

- There is sufficient evidence to suggest that at least one slope is non-zero.

:::

</br> 

#### 1d

Which predictors, if any, are significant predictors of systolic blood pressure? Test at the $\alpha=0.05$ level. Remember to typeset the results.

::: {.panel-tabset}

## Code

```{r}
#| eval: false
c1 <- data.frame(round(summary(m1)$coeff, 3)) # m1 coefficients rounded to 3 decimal places

colnames(c1)[c(2,4)] <- c("Std.Error", "p.value") # re-named 2 columns

formattable(
  c1, 
  list(
    t.value = color_tile("#FEFEB0", "#FEFEB0"),
    p.value = color_tile("#CDFFE0", "#CDFFE0")
    )
  )
```

## Output

```{r}
#| echo: false
c1 <- data.frame(round(summary(m1)$coeff, 3)) # m1 coefficients rounded to 3 decimal places

colnames(c1)[c(2,4)] <- c("Std.Error", "p.value") # re-named 2 columns

formattable(
  c1, 
  list(
    t.value = color_tile("#FEFEB0", "#FEFEB0"),
    p.value = color_tile("#CDFFE0", "#CDFFE0")
    )
  )
```

## Results: age

- Hypotheses:

  - $H_{0}$ : $\beta_{age} = 0$

  - $H_{1}$ : $\beta_{age} \ne 0$

- Test Statistic:

  - $t_{0} = 17.480$

- P-Value:

  - $p < 0.001$

- Rejection Region:

  - Reject $H_{0}$ if $p < \alpha;$ $\alpha = 0.05$

- Conclusion:

  - Reject $H_{0}.$ There is sufficient evidence to suggest that age significantly predicts systolic blood pressure (sbp).

## Results: HSgrad

- Hypotheses:

  - $H_{0}$ : $\beta_{HSgrad} = 0$

  - $H_{1}$ : $\beta_{HSgrad} \ne 0$

- Test Statistic:

  - $t_{0} = -0.983$

- P-Value:

  - $p = 0.326$

- Rejection Region:

  - Reject $H_{0}$ if $p < \alpha;$ $\alpha = 0.05$

- Conclusion:

  - Fail to reject $H_{0}.$ There is not sufficient evidence to suggest that High School Graduate (HSgrad) significantly predicts systolic blood pressure (sbp).

## Results: BMI

- Hypotheses:

  - $H_{0}$ : $\beta_{BMI} = 0$

  - $H_{1}$ : $\beta_{BMI} \ne 0$

- Test Statistic:

  - $t_{0} = 5.870$

- P-Value:

  - $p < 0.001$

- Rejection Region:

  - Reject $H_{0}$ if $p < \alpha;$ $\alpha = 0.05$

- Conclusion:

  - Reject $H_{0}.$ There is sufficient evidence to suggest that Body Mass Index (BMI) significantly predicts systolic blood pressure (sbp).
  
:::

</br>

#### 1e

Provide brief interpretations for the slopes of the predictors.

::: {.panel-tabset}

## Results

- For every $1\phantom/year$ increase in age, systolic blood pressure increases by $0.44\phantom/mmHg$

- Being a high school graduate, decreases systolic blood pressure by $0.84\phantom/mmHg$

- For every $1\phantom/kg/m^{2}$ increase in BMI, systolic blood pressure increases by $0.24\phantom/mmHg$

:::

</br>

#### 1f

Construct an appropriate data visualization to help with explaining the model results. Systolic blood pressure should be on the y-axis, age should be on the x-axis. You choose what lines to create.

::: {.panel-tabset}

## Code

```{r}
#| eval: false
JHS1 <- JHS1 %>% 
  mutate(y_hat_HSgrad.NO_BMI.20 = c1[1,1] + c1[2,1]*age + c1[3,1]*0 + c1[4,1]*20,
         y_hat_HSgrad.NO_BMI.40 = c1[1,1] + c1[2,1]*age + c1[3,1]*0 + c1[4,1]*40,
         y_hat_HSgrad.NO_BMI.60 = c1[1,1] + c1[2,1]*age + c1[3,1]*0 + c1[4,1]*60)

JHS1$HSgrad.factor <- factor(JHS1$HSgrad, levels = c(0, 1), labels = c("No", "Yes"))

JHS1 %>% 
  ggplot(aes(x = age)) +
  geom_point(aes(y = sbp, shape = HSgrad.factor, fill = HSgrad.factor, color = HSgrad.factor), 
             size = 2, 
             alpha = 0.5) +
  scale_shape_manual(values = c(24, 21)) +
  scale_fill_manual(values = c("plum1", "cornsilk2")) +
  scale_color_manual(values = c("orchid4", "cornsilk4")) +
  geom_line(aes(y = y_hat_HSgrad.NO_BMI.20), color = "darkorchid4", linewidth = 1.5, linetype = 3) +
  geom_line(aes(y = y_hat_HSgrad.NO_BMI.40), color = "darkorchid3", linewidth = 1, linetype = 2) +
  geom_line(aes(y = y_hat_HSgrad.NO_BMI.60), color = "darkorchid1", linewidth = 1, linetype = 1) +
  geom_text(aes(x = 85, y = 137, label = "BMI = 20"), color = "darkorchid4", hjust = "left") +
  geom_text(aes(x = 85, y = 142, label = "BMI = 40"), color = "darkorchid3", hjust = "left") +
  geom_text(aes(x = 85, y = 147, label = "BMI = 60"), color = "darkorchid1", hjust = "left") +
  scale_x_continuous(breaks = seq(20,90,10), limits=c(NA,90)) +
  labs(
    x = "Age", 
    y = "Systolic Blood Pressure", 
    title = expression(paste(hat(sbp)," = 95.04 + 0.44(age) - 0.84(HSgrad = 0) + 0.24(BMI: 20,40,60)")),
    color = "High School Graduate", 
    shape = "High School Graduate", 
    fill = "High School Graduate") +
  ylim(NA, 170) +
  theme_bw() +
  theme(
    title = element_text(size=12, face="bold"),
    legend.position = "top", 
    legend.justification = "left",
    plot.background = element_rect(fill="#F9F8F1"),
    legend.background = element_rect(color="cornsilk4"))
```

## Output

```{r}
#| echo: false
JHS1 <- JHS1 %>% 
  mutate(y_hat_HSgrad.NO_BMI.20 = c1[1,1] + c1[2,1]*age + c1[3,1]*0 + c1[4,1]*20,
         y_hat_HSgrad.NO_BMI.40 = c1[1,1] + c1[2,1]*age + c1[3,1]*0 + c1[4,1]*40,
         y_hat_HSgrad.NO_BMI.60 = c1[1,1] + c1[2,1]*age + c1[3,1]*0 + c1[4,1]*60)

JHS1$HSgrad.factor <- factor(JHS1$HSgrad, levels = c(0, 1), labels = c("No", "Yes"))

JHS1 %>% 
  ggplot(aes(x = age)) +
  geom_point(aes(y = sbp, shape = HSgrad.factor, fill = HSgrad.factor, color = HSgrad.factor), 
             size = 2, 
             alpha = 0.5) +
  scale_shape_manual(values = c(24, 21)) +
  scale_fill_manual(values = c("plum1", "cornsilk2")) +
  scale_color_manual(values = c("orchid4", "cornsilk4")) +
  geom_line(aes(y = y_hat_HSgrad.NO_BMI.20), color = "darkorchid4", linewidth = 1.5, linetype = 3) +
  geom_line(aes(y = y_hat_HSgrad.NO_BMI.40), color = "darkorchid3", linewidth = 1, linetype = 2) +
  geom_line(aes(y = y_hat_HSgrad.NO_BMI.60), color = "darkorchid1", linewidth = 1, linetype = 1) +
  geom_text(aes(x = 85, y = 137, label = "BMI = 20"), color = "darkorchid4", hjust = "left") +
  geom_text(aes(x = 85, y = 142, label = "BMI = 40"), color = "darkorchid3", hjust = "left") +
  geom_text(aes(x = 85, y = 147, label = "BMI = 60"), color = "darkorchid1", hjust = "left") +
  scale_x_continuous(breaks = seq(20,90,10), limits=c(NA,90)) +
  labs(
    x = "Age", 
    y = "Systolic Blood Pressure", 
    title = expression(paste(hat(sbp)," = 95.04 + 0.44(age) - 0.84(HSgrad = 0) + 0.24(BMI: 20,40,60)")),
    color = "High School Graduate", 
    shape = "High School Graduate", 
    fill = "High School Graduate") +
  ylim(NA, 170) +
  theme_bw() +
  theme(
    title = element_text(size=12, face="bold"),
    legend.position = "top", 
    legend.justification = "left",
    plot.background = element_rect(fill="#F9F8F1"),
    legend.background = element_rect(color="cornsilk4"))
```

:::

</br>

#### 1g

How many suspected outliers exist? You must justify your answer statistically.

</br>

#### 1h

How many suspected influential/leverage points exist? You must justify your answer statistically.

</br>

#### 1i

Is multicollinearity a problem in this model? You must justify your answer statistically.

</br>

#### 1j

Assess the assumptions on the linear model. Remember to draw your conclusion with appropriate justification.

</br>



###

#### 2a

Model systolic blood pressure (*sbp*; mmHg) as a function of age (*age*; years), education (*HSgrad*; 0=no, 1=yes), and health status as defined by body mass index (*BMI3cat*; 0=poor health, 1=intermediate health, 2=ideal health). Remember to report the resulting model.

#### 2b

Construct the 95% confidence intervals for the regression coefficients. Remember to state the resulting confidence intervals.

#### 2c

Which predictors, if any, are significant predictors of systolic blood pressure? Test at the $\alpha=0.05$ level. Remember to typeset the results.

#### 2d

Provide brief interpretations for the slopes of the predictors.

#### 2e

Construct an appropriate data visualization to help with explaining the model results. Systolic blood pressure should be on the y-axis, age should be on the x-axis. You choose what lines to create.

</br>

###

Required for graduate students / extra credit for undergraduate students: Write a paragraph to summarize the above analysis, written such that a non-quantitative person could understand. Include if you recommend using BMI as a continuous variable or a categorical variable. For full credit, you must explain your reasoning.

